---
title: "Effect of Base Salary on Overtime"
author: "Shaun Stearns"
date: "8/21/2019"
output: word_document
---

#Load Necessary Libraries
```{r}
library(psych)
library(QuantPsyc)
library(mixlm)
library(ez)
library(readr)
library(dplyr)
library(stats)
library(DescTools)
library(ggplot2)
library(ggthemes)
library(extrafont)
library(jtools)
```


#Set directory
```{r}
setwd("Desktop/CSU Global Data Analytics/MIS500/Portfolio Project") 
```


#Load file and replace 0 with NA
```{r}
NYC_Fisc_Clean <- read_csv("NYC_Fisc_Clean.csv", 
                           col_types = cols(`Base Salary` = col_number(), 
                                            `OT Hours` = col_number(), `Regular Gross Paid` = col_number(), 
                                            `Regular Hours` = col_number(), `Total OT Paid` = col_number(), 
                                            `Total Other Pay` = col_number()),
                                              na="0")
```

#Check File
###View(NYC_Fisc_Clean)
###Descriptives for Borough and Over Time Hours. Determine Boroughs with the highest sample.
```{r}
describeBy(NYC_Fisc_Clean$`OT Hours`, NYC_Fisc_Clean$`Work Location Borough`)
```


#Remove Rows with NA Values for Over Time Hours, Base Salary, and Total Other Pay
```{r}
NYC_Fisc_Clean<-NYC_Fisc_Clean[complete.cases(NYC_Fisc_Clean$`OT Hours`),]
NYC_Fisc_Clean<-NYC_Fisc_Clean[complete.cases(NYC_Fisc_Clean$`Base Salary`),]
NYC_Fisc_Clean<-NYC_Fisc_Clean[complete.cases(NYC_Fisc_Clean$`Total Other Pay`),]
```

#Split file based on Borough Location
```{r}
NYC_Split <- split(NYC_Fisc_Clean, NYC_Fisc_Clean$`Work Location Borough`)
```

#Create datasets based on Borough location with the largest sample sizes
```{r}
BRONX<-NYC_Split[[2]]
BROOKLYN<-NYC_Split[[3]]
MANHATTAN<-NYC_Split[[7]]
QUEENS<-NYC_Split[[11]]
RICHMOND<-NYC_Split[[12]]
WESTCHESTER<-NYC_Split[[16]]
```


```{r}
#Combine Split files into a single file with chosen Boroughs
BigBoroughs <- rbind(BRONX, BROOKLYN, MANHATTAN, QUEENS, RICHMOND, WESTCHESTER)
#Overtime Hours
describeBy(BigBoroughs$`OT Hours`, BigBoroughs$`Work Location Borough`)
#Base Salary
describeBy(BigBoroughs$`Base Salary`, BigBoroughs$`Work Location Borough`)
#Total Other Pay
describeBy(BigBoroughs$`Total Other Pay`, BigBoroughs$`Work Location Borough`)
```

#Create Non-transformed dataset for visualizations
```{r}
BigBoroughs2 <- BigBoroughs
#View(BigBuoroughs2)
```


#Log Transform to normalize data
```{r}
BigBoroughs$`OT Hours` = log10(BigBoroughs$`OT Hours`)
BigBoroughs$`Base Salary` = log10(BigBoroughs$`Base Salary`)
BigBoroughs$`Total Other Pay` = log10(BigBoroughs$`Total Other Pay`)
View(BigBoroughs)
```


#Analysis of Variance with Unadjusted and adjusted post hoc tests
```{r}
anova1 <- aov(BigBoroughs$`OT Hours` ~ BigBoroughs$`Work Location Borough`)
summary(anova1, digits = 4)
#Post hoc unadjusted: Similar to a t-test but takes into account overall error
Post<-TukeyHSD(anova1)
Post
#Scheffe Post Hoc: Statistical adjustment based on multiple comparisons
PostSchf <- ScheffeTest(anova1, conf.level=NA)
PostSchf
```

```{r}
anova2 <- aov(BigBoroughs$`Base Salary` ~ BigBoroughs$`Work Location Borough`)
summary(anova2)
#Post hoc unadjusted: Similar to a t-test but takes into account overall error
Post<-TukeyHSD(anova2)
Post
#Scheffe Post Hoc: Statistical adjustment based on multiple comparisons
PostSchf <- ScheffeTest(anova2, conf.level=NA)
PostSchf
```

```{r}
anova3 <- aov(BigBoroughs$`Total Other Pay` ~ BigBoroughs$`Work Location Borough`)
summary(anova3)
#Post hoc unadjusted: Similar to a t-test but takes into account overall error
Post<-TukeyHSD(anova3)
Post
#Scheffe Post Hoc: Statistical adjustment based on multiple comparisons
PostSchf <- ScheffeTest(anova3, conf.level=NA)
PostSchf
```


#Multiple regression on split files to determine what variables predict increased overtime
```{r}
#BRONX
BRONXlm<-lm(BRONX$`OT Hours` ~ BRONX$`Base Salary`)
BRONXmlm<-lm(BRONX$`OT Hours` ~ BRONX$`Base Salary` +
               BRONX$`Total Other Pay`, data = BRONX)
summ(BRONXlm, digits = 4)
lm.beta(BRONXlm)
summ(BRONXmlm, digits = 4)
lm.beta(BRONXmlm)
```

```{r}
#BROOKLYN
BROOKLYNlm<-lm(BROOKLYN$`OT Hours` ~ BROOKLYN$`Base Salary`)
BROOKLYNmlm<-lm(BROOKLYN$`OT Hours` ~ BROOKLYN$`Base Salary` +
                  BROOKLYN$`Total Other Pay`, data = BROOKLYN)
summ(BROOKLYNlm, digits = 4)
lm.beta(BROOKLYNlm)
summ(BROOKLYNmlm, digits = 4)
lm.beta(BROOKLYNmlm)
```

```{r}
#MANHATTAN
MANHATTANlm<-lm(MANHATTAN$`OT Hours` ~ MANHATTAN$`Base Salary`)
MANHATTANmlm<-lm(MANHATTAN$`OT Hours` ~ MANHATTAN$`Base Salary` +
                  MANHATTAN$`Total Other Pay`, data = MANHATTAN)
summ(MANHATTANlm, digits = 4)
lm.beta(MANHATTANlm)
summ(MANHATTANmlm, digits = 4)
lm.beta(MANHATTANmlm)
```

```{r}
#QUEENS
QUEENSlm<-lm(QUEENS$`OT Hours` ~ QUEENS$`Base Salary`)
QUEENSmlm<-lm(QUEENS$`OT Hours` ~ QUEENS$`Base Salary` +
                   QUEENS$`Total Other Pay`, data = QUEENS)
summ(QUEENSlm, digits = 4)
lm.beta(QUEENSlm)
summ(QUEENSmlm, digits = 4)
lm.beta(QUEENSmlm)
```

```{r}
#RICHMOND
RICHMONDlm<-lm(RICHMOND$`OT Hours` ~ RICHMOND$`Base Salary`)
RICHMONDmlm<-lm(RICHMOND$`OT Hours` ~ RICHMOND$`Base Salary` +
                RICHMOND$`Total Other Pay`, data = RICHMOND)
summ(RICHMONDlm, digits = 4)
lm.beta(RICHMONDlm)
summ(RICHMONDmlm, digits = 4)
lm.beta(RICHMONDmlm)
```

```{r}
#WESTCHESTER
WESTCHESTERlm<-lm(WESTCHESTER$`OT Hours` ~ WESTCHESTER$`Base Salary`)
WESTCHESTERmlm<-lm(WESTCHESTER$`OT Hours` ~ WESTCHESTER$`Base Salary` +
                  WESTCHESTER$`Total Other Pay`, data = WESTCHESTER)
summ(WESTCHESTERlm, digits = 4)
lm.beta(WESTCHESTERlm)
summ(WESTCHESTERmlm,digits = 4)
lm.beta(WESTCHESTERmlm)
```


#Plots: Bar Chart (Overtime by Boroughs)
```{r}
A <- ggplot(data=BigBoroughs2, aes(x=reorder(BigBoroughs2$`Work Location Borough`, BigBoroughs2$`OT Hours`),
                             y=BigBoroughs2$`OT Hours`, fill=`Work Location Borough`)) +
  geom_bar(stat="summary", fun.y = "mean") + 
  xlab("NYC Borough") + 
  ylab("Hours of Overtime") +
  coord_cartesian(ylim = c(100, 300)) +
  ggtitle("NYC Employee Overtime by Borough")
A + scale_y_continuous(labels = scales::comma) + 
  theme(axis.text.x =
            element_text(size  = 10, angle = 45, hjust = 1, vjust = 1)) +
  guides(fill=guide_legend(title=NULL))
```


#Plots: Bar Chart (Basesalary by Borough)
```{r}
B <- ggplot(data=BigBoroughs2, aes(x=reorder(BigBoroughs2$`Work Location Borough`, BigBoroughs2$`Base Salary`), 
                             y=BigBoroughs2$`Base Salary`, fill=`Work Location Borough`)) +
  geom_bar(stat="summary", fun.y = "mean", aes(fill=`Work Location Borough`)) +
  xlab("NYC Borough") + 
  ylab("Base Salary") + 
  coord_cartesian(ylim = c(55000, 62000)) +
  ggtitle("NYC Employee Base Salary by Borough")
B + scale_y_continuous(labels = scales::comma) +
  theme(axis.text.x =
            element_text(size  = 10, angle = 45, hjust = 1, vjust = 1)) +
  guides(fill=guide_legend(title=NULL))
```


#Plots: Bar Chart (Other Pay by Borough)
```{r}
C <- ggplot(data=BigBoroughs2, aes(x=reorder(BigBoroughs2$`Work Location Borough`, BigBoroughs2$`Total Other Pay`), 
                             y=BigBoroughs2$`Total Other Pay`, fill=`Work Location Borough`)) +
  geom_bar(stat="summary", fun.y = "mean", aes(fill=`Work Location Borough`)) +
  xlab("NYC Borough") + 
  ylab("Total Other Pay") + 
  coord_cartesian(ylim = c(4000, 8300)) +
  ggtitle("NYC Employee Total Other Pay by Borough")
C + scale_y_continuous(labels = scales::comma) + 
  theme(axis.text.x =
            element_text(size  = 10, angle = 45, hjust = 1, vjust = 1)) +
  guides(fill=guide_legend(title=NULL))
```
